---
title: "EDA-Covid-Data"
author: "Alexander Mars"
date: "30/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Cases Data
## Cases by location
Getting cases by location and creating a count per LGA for each date, may be useful for visualising number of cases vs. LGA trends over time. 
```{r}
library(dplyr)
# Contains the date and LGA of each case
# can aggregate by LGA for each date to get a count by LGA
cases.location <- read.csv("https://data.nsw.gov.au/data/dataset/aefcde60-3b0c-4bc0-9af1-6fe652944ec2/resource/21304414-1ff1-4243-a5d2-f52778048b29/download/confirmed_cases_table1_location.csv", header = T)
head(cases.location)
dim(cases.location)
# retain relevant columns
keep <- c("notification_date", "postcode", "lga_name19")
cases.location <- cases.location[, names(cases.location) %in% keep]
# Missing values
colSums(cases.location == "None") # there are 1043 missing postcode values
colSums(cases.location == "") # there are 1043 missing location values total
cases.location[cases.location ==""] <- NA
cases.location[cases.location == "None"] <- NA
colSums(is.na(cases.location))
# Drop rows with missing values
cases.location.complete <- na.omit(cases.location)
# Save complete data
write.csv(cases.location.complete, "data/cases/processed/confirmed_cases_location_complete.csv", 
          row.names = FALSE)
# Count the number of cases per postcode for each date
cases.location.count <- cases.location.complete %>%
  group_by(notification_date, postcode) %>%
  summarise(n = n())
dim(cases.location.count)
colnames(cases.location.count) <- c("date", "postcode", "cases")
tail(cases.location.count)
# Save output
write.csv(cases.location.count, "data/cases/processed/confirmed_cases_count_by_postcode.csv", 
          row.names = FALSE)
# Count the number of cases per LGA for each date
cases.location.count.lga <- cases.location.complete %>%
  group_by(notification_date, lga_name19) %>%
  summarise(n = n())
dim(cases.location.count.lga)
colnames(cases.location.count.lga) <- c("date", "lga", "cases")
tail(cases.location.count.lga)
# Save output
write.csv(cases.location.count.lga, "data/cases/processed/confirmed_cases_count_by_lga.csv", 
          row.names = FALSE)

# Total number of cases per day
cases.day.count <- cases.location %>%
  group_by(notification_date) %>%
  summarise(cases = n())
colnames(cases.day.count) <- c("date", "cases")
head(cases.day.count)
# save output
write.csv(cases.day.count, "data/cases/processed/confirmed_cases_per_day.csv", 
          row.names = FALSE)
# Inspect time period
min(cases.day.count$date)
max(cases.day.count$date)
max(as.Date(cases.day.count$date)) - min(as.Date(cases.day.count$date))
```
## Cases by age group
The age group data doesn't appear to match up to the location data above, 
it covers a shorter time period
```{r}
# Loading in the age data
cases.age <- read.csv("data/cases/confirmed_cases_table2_age_group.csv", header = T)
head(cases.age)
dim(cases.age) # observations don't match up to the confirmed cases table
# Processing
# Convert to date format
cases.age$notification_date <- as.Date(cases.age$notification_date)
# Remove "AgeGroup_" from age_group column
cases.age$age_group <- gsub("AgeGroup_", "", cases.age$age_group)
# Recheck for missing values
sum(cases.age$age_group == "None")
cases.age[cases.age$age_group == "None",] <- NA
colSums(is.na(cases.age))
# Drop missing values
cases.age.complete <- na.omit(cases.age)
# Date range
max(cases.age.complete$notification_date)-min(cases.age.complete$notification_date)
min(cases.age.complete$notification_date) # starting in june 2021
# Save output
write.csv(cases.age.complete, "data/cases/processed/confirmed_cases_age_complete.csv", 
          row.names = FALSE)
```

## Cases by likely source

Nothing was done with the likely source data alone, but the source + location data
was processed.
```{r}
# Can process the data to have a local/overseas column, and a linked/unlinked column
cases.source.location <- read.csv("https://data.nsw.gov.au/data/dataset/97ea2424-abaf-4f3e-a9f2-b5c883f42b6a/resource/2776dbb8-f807-4fb2-b1ed-184a6fc2c8aa/download/confirmed_cases_table4_location_likely_source.csv", header = T)
head(cases.source.location, 10)
# Missing values
cases.source.location[cases.source.location ==""] <- NA
cases.source.location[cases.source.location == "None"] <- NA
colSums(is.na(cases.source.location))
cases.source.location.complete <- na.omit(cases.source.location)

# Split the source into local/overseas and known/unknown
# create local, overseas, interstate column
unique(cases.source.location.complete$likely_source_of_infection)
source <- strsplit(cases.source.location.complete$likely_source_of_infection, " ")
source <- lapply(source, function(split.list) split.list[1])
source <- unlist(source)
cases.source.location.complete$source <- source
# create known/unknown column
# applies only to locally acquired cases
# 0 means not local, 1 means unknown, 2 means ongoing, 3 means known
known <- strsplit(cases.source.location.complete$likely_source_of_infection, "-")
known <- lapply(known, function(split.list) split.list[2])
known[is.na(known)] <- 0
known[known == " no links to known case or cluster"] <- 1
known[known == " investigation ongoing"] <- 2
known[known == " linked to known case or cluster"] <- 3
known <- unlist(known)
cases.source.location.complete$known <- known
head(cases.source.location.complete)
keepList <- c("notification_date", "source", "known", "postcode", "lga_name19")

cases.source.location.complete = subset(cases.source.location.complete, 
                                        select = keepList)
head(cases.source.location.complete)
# save output
write.csv(cases.source.location.complete, "data/cases/processed/confirmed_cases_source_location.csv", 
          row.names = FALSE)
```


# Tests Data
Only the aggregated data was dealt with (unaggregated seems useless)
```{r}
# Aggregated Data
test.pcr.aggregated <- read.csv("data/tests/pcr_testing_table2_age_group_agg.csv",
                                header = T)
str(test.pcr.aggregated)
# Replace "AgeGroup_" with ""
test.pcr.aggregated$age_group <- gsub("AgeGroup_", "", test.pcr.aggregated$age_group)
# Check for missing values
unique(test.pcr.aggregated$age_group) 
sum(test.pcr.aggregated$age_group == "None") # 139 instances without an age group
test.pcr.aggregated$test_count <- as.numeric(test.pcr.aggregated$test_count)
head(test.pcr.aggregated)
test.pcr.aggregated[test.pcr.aggregated == "None"] <- NA
test.pcr.aggregated.complete <- na.omit(test.pcr.aggregated)
# Save completed data
write.csv(test.pcr.aggregated.complete, "data/tests/processed/pcr_testing_age_group_complete.csv", 
          row.names = FALSE)
min(test.pcr.aggregated.complete$test_date)
max(test.pcr.aggregated.complete$test_date)
# Save data with good age labels
write.csv(test.pcr.aggregated, "data/tests/processed/pcr_testing_age_group.csv", 
          row.names = FALSE)
```

Create a total tests by date table (sum over all age groups)
```{r}
pcr.test.count <- test.pcr.aggregated %>%
  group_by(test_date) %>%
  summarise(tests = sum(test_count))
tail(pcr.test.count)
write.csv(pcr.test.count, "data/tests/processed/total_tests_by_date.csv", 
          row.names = FALSE)
```

Testing by location data:
```{r}
pcr.location <- read.csv("data/tests/pcr_testing_table1_location_agg.csv", header = T)
colSums(pcr.location == "None")
pcr.location[pcr.location == "None"] <- NA
pcr.location.complete <- na.omit(pcr.location)
pcr.location.complete$test_date <- as.Date(pcr.location.complete$test_date)
min(pcr.location.complete$test_date)
max(pcr.location.complete$test_date)
max(pcr.location.complete$test_date) - min(pcr.location.complete$test_date)
head(pcr.location.complete)
# Save output
write.csv(pcr.location.complete, "data/tests/processed/test_by_location_complete.csv", 
          row.names = FALSE)
```

# Clinics Data
Seeing as postcode is used for test and cases data, 
we could inspect the number of clinics per postcode as an indicator 
of testing availability.
```{r}
clinics <- read.csv("data/clinics/nsw-health-covid-19-test-clinics-2021-09-30-0915.csv", 
                    header = T)
str(clinics)
clinics.by.postcode <- clinics[names(clinics) %in% c("title", "Postcode")]
clinics.by.postcode.count <- clinics.by.postcode %>%
  group_by(Postcode) %>%
  summarise(clinics = n())
# save output
write.csv(clinics.by.postcode.count, "data/clinics/processed/num_clinics_by_postcode.csv", 
          row.names = FALSE)
head(clinics.by.postcode.count)
range(clinics.by.postcode.count$clinics) # not a huge difference anyway
```

Transport and flight data not included.

