---
title: "NSW Geo Data"
author: 'UniKey: amar6958, SID:450347582'
date: "05/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read Shapefile

```{r}
library(rgdal)
nsw.shp <- readOGR( 
  dsn="data/nswGeo/LocalGovernmentArea.shp",  verbose=FALSE
)
```
# Link cases by LGA to the shapefile

We can link the datasets by making sure the LGA names match up. 
```{r}
cases.count.by.lga <- read.csv("data/cases/processed/confirmed_cases_count_by_lga.csv", header = T)
unique(cases.count.by.lga$lga)
# To make the lga names in the cases data match the geo data lga names,
# we need to capitalise them and remove the characters in brackets
cases.count.by.lga$lga <- unlist(lapply(cases.count.by.lga$lga, 
                                        function (lga) {
                                          toupper(gsub(" \\(.*?\\)", "", lga))
                                          }))

```
## Checking for differences

There may be differences, due to different naming conventions for the LGAs, or
due to the fact that some LGAs may not have had any cases during the pandemic. 

```{r}
nsw.shp$lganame[!(nsw.shp$lganame %in% cases.count.by.lga$lga)]
```
```{r}
unique(cases.count.by.lga$lga[!(cases.count.by.lga$lga %in% nsw.shp$lganame)])
```
We can adjust these manually, except for correctional settings, which 
appears to be a count of COVID cases that have occurred in Prison. Since this 
does not relate to geographic information and is not impacted by mobility or 
lockdown measures, we should probably remove information from correctional 
settings. The rest of the suburb names can be corrected manually.
```{r}
cases.count.by.lga[cases.count.by.lga$lga == "PARRAMATTA", "lga"] <- "CITY OF PARRAMATTA"
cases.count.by.lga[cases.count.by.lga$lga == "LITHGOW", "lga"] <- "LITHGOW CITY"
cases.count.by.lga[cases.count.by.lga$lga == "NAMBUCCA", "lga"] <- "NAMBUCCA VALLEY"
cases.count.by.lga[cases.count.by.lga$lga == "ALBURY", "lga"] <- "ALBURY CITY"
cases.count.by.lga[cases.count.by.lga$lga == "UPPER HUNTER SHIRE", "lga"] <- "UPPER HUNTER"
cases.count.by.lga[cases.count.by.lga$lga == "WARRUMBUNGLE SHIRE", "lga"] <- "WARRUMBUNGLE"
cases.count.by.lga.reduced <- cases.count.by.lga[cases.count.by.lga$lga != "CORRECTIONAL SETTINGS", ]
cases.count.by.lga.reduced
unique(cases.count.by.lga.reduced$lga[!(cases.count.by.lga.reduced$lga %in% nsw.shp$lganame)])
write.csv(cases.count.by.lga.reduced, "data/nswGeo/confirmed_cases_count_by_lga_geo.csv", 
          row.names = FALSE)
```

# Linking Source of Case to Shapefile
We can also link data for known source of cases to the shapefile:
```{r}
cases.source <- read.csv("data/cases/processed/confirmed_cases_source_location.csv", header = 
T)
head(cases.source)
colnames(cases.source) <- c("date", "source", "known", "postcode", "lga")
cases.source$lga <- unlist(lapply(cases.source$lga,
                                  function (lga) {
                                    toupper(gsub(" \\(.*?\\)", "", lga))
                                    }))
```

```{r}
nsw.shp$lganame[!(nsw.shp$lganame %in% cases.source$lga)]
```

```{r}
unique(cases.source$lga[!(cases.source$lga %in% nsw.shp$lganame)])
```
```{r}
cases.source[cases.source$lga == "PARRAMATTA", "lga"] <- "CITY OF PARRAMATTA"
cases.source[cases.source$lga == "LITHGOW", "lga"] <- "LITHGOW CITY"
cases.source[cases.source$lga == "NAMBUCCA", "lga"] <- "NAMBUCCA VALLEY"
cases.source[cases.source$lga == "ALBURY", "lga"] <- "ALBURY CITY"
cases.source[cases.source$lga == "UPPER HUNTER SHIRE", "lga"] <- "UPPER HUNTER"
cases.source[cases.source$lga == "WARRUMBUNGLE SHIRE", "lga"] <- "WARRUMBUNGLE"
cases.source.reduced <- cases.source[cases.source$lga != "CORRECTIONAL SETTINGS", ]
unique(cases.source.reduced$lga[!(cases.source.reduced$lga %in% nsw.shp$lganame)])
write.csv(cases.source.reduced, "data/nswGeo/confirmed_cases_source_by_lga_geo.csv", 
          row.names = FALSE)
```
We can create a count of the number of cases in total, and the count for the different source types, for each LGA on each date.
```{r}
library(dplyr)
head(cases.source.reduced)
cases.source.reduced2 <- cases.source.reduced %>%
  group_by(lga, date, source) %>%
  summarise(cases = n())
cases.source.reduced2
write.csv(cases.source.reduced2, "data/nswGeo/confirmed_cases_source_count_by_lga_geo.csv", 
          row.names = FALSE)
```

# Linking Test Data to Shapefile
```{r}
pcr.location <- read.csv("data/tests/processed/test_by_location_complete.csv", 
          header = T)
head(pcr.location)
pcr.location.shp <- pcr.location[,names(pcr.location) %in% c("test_date",
                                                             "postcode",
                                                             "lga_name19",
                                                             "test_count")]
colnames(pcr.location.shp) <- c("date", "postcode", "lga", "test_count")
head(pcr.location.shp)
pcr.location.shp$lga <- unlist(lapply(pcr.location.shp$lga,
                                  function (lga) {
                                    toupper(gsub(" \\(.*?\\)", "", lga))
                                    }))
```
```{r}
nsw.shp$lganame[!(nsw.shp$lganame %in% pcr.location.shp$lga)]
```
```{r}
unique(pcr.location.shp$lga[!(pcr.location.shp$lga %in% nsw.shp$lganame)])
```
```{r}
pcr.location.shp[pcr.location.shp$lga == "PARRAMATTA", "lga"] <- "CITY OF PARRAMATTA"
pcr.location.shp[pcr.location.shp$lga == "LITHGOW", "lga"] <- "LITHGOW CITY"
pcr.location.shp[pcr.location.shp$lga == "NAMBUCCA", "lga"] <- "NAMBUCCA VALLEY"
pcr.location.shp[pcr.location.shp$lga == "ALBURY", "lga"] <- "ALBURY CITY"
pcr.location.shp[pcr.location.shp$lga == "UPPER HUNTER SHIRE", "lga"] <- "UPPER HUNTER"
pcr.location.shp[pcr.location.shp$lga == "WARRUMBUNGLE SHIRE", "lga"] <- "WARRUMBUNGLE"
pcr.location.shp.reduced <- pcr.location.shp[pcr.location.shp$lga != "CORRECTIONAL SETTINGS", ]
unique(pcr.location.shp.reduced$lga[!(pcr.location.shp.reduced$lga %in% nsw.shp$lganame)])
write.csv(pcr.location.shp.reduced, "data/nswGeo/tests_by_lga_geo.csv", 
          row.names = FALSE)
head(pcr.location.shp.reduced)
```
```{r}
library(dplyr)
head(pcr.location.shp.reduced)
pcr.location.shp.reduced2 <- pcr.location.shp.reduced %>%
  group_by(date, lga) %>%
  summarise(tests = sum(test_count))
median(pcr.location.shp.reduced2$tests)
```
# Link data to LGA populations
```{r}
pop.dat <- read.csv("data/nswGeo/lga_population.csv", header = T)
pop.dat <- pop.dat[ , colSums(is.na(pop.dat)) == 0]
pop.dat$population <- as.numeric(gsub(",", "", pop.dat$population))
pop.dat$density <- as.numeric(pop.dat$density)
pop.dat$lga <- toupper(gsub("Council", "", pop.dat$lga))
pop.dat$lga
move_substr_to_front <- function(string)
{
  splits <- unlist(strsplit(string, ", "))
  if (length(splits) > 1){
  front <- splits[2]
  back <- splits[1]
  return(paste(front, back))
  } else {
    return(string)
  }
}

pop.dat$lga2 <- unlist(lapply(pop.dat$lga, move_substr_to_front))
pop.dat$lga2 <- gsub("SHIRE", "", pop.dat$lga2)
pop.dat$lga2 <- trimws(pop.dat$lga2)
pop.dat$lga2 <- gsub("CITY OF ", "", pop.dat$lga2)
pop.dat$lga2 <- gsub("MUNICIPALITY OF ", "", pop.dat$lga2)
pop.dat
pop.dat[pop.dat$lga2 == "PARRAMATTA", "lga2"] <- "CITY OF PARRAMATTA"
pop.dat[pop.dat$lga2 == "LITHGOW", "lga2"] <- "LITHGOW CITY"
pop.dat[pop.dat$lga2 == "NAMBUCCA", "lga2"] <- "NAMBUCCA VALLEY"
pop.dat[pop.dat$lga2 == "ALBURY", "lga2"] <- "ALBURY CITY"
pop.dat[pop.dat$lga2 == "SUTHERLAND", "lga2"] <- "SUTHERLAND SHIRE"
pop.dat[pop.dat$lga2 == "THE HILLS", "lga2"] <- "THE HILLS SHIRE"
pop.dat[pop.dat$lga2 == "UPPER LACHLAN", "lga2"] <- "UPPER LACHLAN SHIRE"
pop.dat[pop.dat$lga2 == "GREATER HUME", "lga2"] <- "GREATER HUME SHIRE"
pop.dat[pop.dat$lga2 == "BATHURST REGION", "lga2"] <- "BATHURST REGIONAL"
pop.dat[pop.dat$lga2 == "UNINCORPORATED FAR WEST", "lga2"] <- "UNINCORPORATED - FAR WEST AREA"
pop.dat[pop.dat$lga2 == "LORD HOWE ISLAND", "lga2"] <- "UNINCORPORATED - LORD HOWE ISLAND"
pop.dat[pop.dat$lga2 == "QUEANBEYAN–PALERANG REGIONAL", "lga2"] <- "QUEANBEYAN-PALERANG REGIONAL"
pop.dat$lga <- pop.dat$lga2
pop.dat <- pop.dat[, !(names(pop.dat) %in% c("lga2"))]
```
```{r}
unique(pop.dat$lga2[!(pop.dat$lga2 %in% nsw.shp$lganame)])
```


```{r}
nsw.shp$lganame[!(nsw.shp$lganame %in% pop.dat$lga2)]
```
```{r}
pop.dat # Population by LGA
pcr.location.shp.reduced2 # tests by LGA over date
cases.count.by.lga.reduced

pcr.location.shp.reduced.pop <- merge(pcr.location.shp.reduced2, pop.dat, by = "lga")
pcr.location.shp.reduced.pop <- pcr.location.shp.reduced.pop[order(pcr.location.shp.reduced.pop$date),]
pcr.location.shp.reduced.pop$test_pop <- log(pcr.location.shp.reduced.pop$tests / (pcr.location.shp.reduced.pop$population))
normalize <- function(vals, upper = 100)
{
  vals <- (vals - min(vals)) / (max(vals)-min(vals))
  return(round(vals*100, 2)) 
}
pcr.location.shp.reduced.pop$test_pop <- normalize(pcr.location.shp.reduced.pop$test_pop)

cases.count.by.lga.reduced.pop <- merge(cases.count.by.lga.reduced, pop.dat, by = "lga")
cases.count.by.lga.reduced.pop <- cases.count.by.lga.reduced.pop[order(cases.count.by.lga.reduced.pop$date),]
cases.count.by.lga.reduced.pop$case_pop <- log(cases.count.by.lga.reduced.pop$cases / (cases.count.by.lga.reduced.pop$population))
cases.count.by.lga.reduced.pop$case_pop <- normalize(cases.count.by.lga.reduced.pop$case_pop)
cases.count.by.lga.reduced.pop
# save files 
write.csv(pcr.location.shp.reduced.pop, "data/nswGeo/tests_by_lga_pop_geo.csv", 
          row.names = FALSE)
write.csv(cases.count.by.lga.reduced.pop, "data/nswGeo/cases_by_lga_pop_geo.csv", 
          row.names = FALSE)
```
```{r}

```


# Get NSW Postal Areas 

Clinic data is only available by postcode, so we can't use the LGA shapefile
as a postcode may span multiple LGAs. 
```{r}
library(rgdal)
aus.postal <- readOGR( 
  dsn="data/ausGeoPostal/POA_2016_AUST.shp",  verbose=FALSE
)
```

